---
title: "02-Format_Data"
format: html
---

This file formats all raw data used in this project in combination with the formatted municipality, canton, entity, and IEBL shapefiles formatted in 01-Format_Shapefiles.qmd. Data is exported as .csv files

## Loading
```{r}
#| label: load-libraries
#| echo: FALSE

library(geodata)
library(units)
library(sf)
library(mapview)
library(readxl)
library(units)
library(tidyverse)

```

```{r}
#| label: load-data
#| echo: FALSE

### CENSUS DATA ###

# Census Data, 1991
# Source: xxx
bih_census_1991 <- readxl::read_xlsx("../Data/bih_census_1991.xlsx")

# Census Data, 1991, by Community
# Source: xxx
bih_census_1991_community <- readxl::read_xlsx("../Data/Stanovništvo prema nacionalnom izjašnjavanju po mjesnim zajednicama 1991.xlsx")

# Census Data, 1991, by Post-War Municipalities
# Source: Avdić, Mirić, Gekić, and Gekić
bih_census_1991_postwar_modified <- readxl::read_xlsx("../Data/Avdić, Mirić, Gekić, and Gekić data.xlsx", skip = 1)

# Census Data, 2013
# Source: xxx
bih_census_2013 <- readxl::read_xlsx("../Data/bih_census_2013.xlsx")


### SHAPE FILES ###

# Pre-War Shapefile
bih_prewar_municipalities_shapefile <- sf::read_sf(
  "../Shape Files/Formatted Shape Files/bih_prewar_municipalities_shapefile_formatted.gpkg",
  quiet = TRUE
)

# Post-War Municipality Shapefile
bih_postwar_municipalities_shapefile <- sf::read_sf(
  "../Shape Files/Formatted Shape Files/bih_postwar_municipalities_shapefile_formatted.gpkg",
  quiet = TRUE
)

# Post-War Canton Shapefile
bih_postwar_cantons_shapefile <- sf::read_sf(
  "../Shape Files/Formatted Shape Files/bih_postwar_cantons_shapefile_formatted.gpkg",
  quiet = TRUE
)

# Post-War Entity Shapefile
bih_postwar_entities_shapefile <- sf::st_read(
  "../Shape Files/Raw Shapefiles/geoBoundaries-BIH-ADM1-all/geoBoundaries-BIH-ADM1.shp",
  quiet = TRUE
  )

# Post-War IEBL Shapefile
bih_postwar_iebl_shapefile <- sf::read_sf(
  "../Shape Files/Formatted Shape Files/iebl_shapefile_formatted.gpkg",
  quiet = TRUE
)

# Country Outline Shapefile
bih_outline_shapefile <- sf::st_read(
  "../Shape Files/Raw Shapefiles/geoBoundaries-BIH-ADM0-all/geoBoundaries-BIH-ADM0.shp",
  quiet = TRUE
  )

# Croatia, Serbia, Montenegro, and Kosovo County Outline Shapefiles
hrv_shapefile <- geodata::gadm(country = "HRV", level = 0)
srb_shapefile <- geodata::gadm(country = "SRB", level = 0)
mne_shapefile <- geodata::gadm(country = "MNE", level = 0)
xko_shapefile <- geodata::gadm(country = "XKO", level = 0)

# Minefield 2004 Map Shapefile
# Source: xxx
minefield_2004_shapefile <- sf::read_sf(
  "../Shape Files/bih_landmine_areas.gpkg",
  quiet = TRUE
)


### ELECTION DATA ###

# Municipal Vote Tally
# Source: xxx
municipal_election_tally <- readxl::read_xlsx("../Data/bih_municipal_votes_97_tally.xlsx")

# Municipal Vote Seat Allocation
# Source: xxx
municipal_seat_allocation <- read.csv("../Data/bih_municipal_elections_97_seat_allocation.csv")

# Nationalist Party Leanings
# Source: xxx
nationalist_party_leanings_1997 <- readxl::read_xlsx("../Data/bih_parties_1997_leanings.xlsx")
 
# Voting Location
# Source: xxx
voting_location_1997 <- readxl::read_xlsx("../Data/bih_voting_location_1997.xlsx")


### OTHER DATA ###
# Death Data, by Post-War Municipalities
# Source: Costalli and Moro, 2012
death_data_cm_original <- readxl::read_xlsx("../Data/deaths_per_municipality.xlsx")

# UCDP GED Event Dataset v25.1
# Source: xxx
gede <- readxl::read_xlsx("../Data/GEDEvent_v25_1.xlsx")

# Post-War Municipalities Metadata
postwar_metadata <- readxl::read_xlsx("../Data/bih_postwar_municipality_metadata.xlsx")

# Municipality Reorganization Crosswalk
municipality_reorg_crosswalk <- readxl::read_xlsx("../Data/bih_municipality_reorg_crosswalk.xlsx")

# Minefield Prevalence by Canton
# Source: xxx
landmines <- readxl::read_xlsx("../Data/unmac_minefield_canton_data_1997.xlsx", skip = 1)

```

## Pre-War Data
This section formats data that is from before the 1992-1995 war or is based upon the pre-war municipalities.

### 1991 Census Data
```{r}
#| label: census-data-1991
#| echo: FALSE

# Note: The total row's Total value shows a population 1,402 individuals greater than the sum of the
# individual ethnic groups. All municipality-level Total values match the sum of the breakdown
# values.
census_data_1991 <- bih_census_1991 %>%
  # drop Bosnia-wide row
  dplyr::filter(Municipality != "Total") %>%
  # rename Muslims (Bosniaks) and Total variables
  dplyr::rename(
    muslims = `Muslims by nationality`,
    population = Total) %>%
  dplyr::rename_with(tolower) %>%
  dplyr::mutate(
    # calculate municipality-level percentages by ethnicity
    dplyr::across(c(muslims, serbs, croats, yugoslavs, other), ~ .x / population,
                  .names = "percent_{.col}"),
    # calculate polarization and fractionalization metrics
    fractionalization = 1 - rowSums(dplyr::across(dplyr::starts_with("percent_"))^2),
    polarization = 1 - 4 * rowSums(dplyr::across(dplyr::starts_with("percent_"), ~ .x * (0.5 - .x)^2)),
    percent_total_population = population / sum(population),
    # rename pre-war municipalities to standardize names
    municipality = dplyr::case_match(
      municipality,
      "Centar Sarajevo" ~ "Sarajevo Centar",
      "Dubica" ~ "Bosanska Dubica",
      "Gradiška" ~ "Bosanska Gradiška",
      "Novi Grad" ~ "Bosanski Novi",
      "Novi Grad Sarajevo" ~ "Sarajevo Novi Grad",
      "Novo Sarajevo" ~ "Sarajevo Novo Sarajevo",
      "Šamac" ~ "Bosanski Šamac",
      "Stari Grad Sarajevo" ~ "Sarajevo Stari Grad",
      "Vogošća" ~ "Sarajevo Vogošća",
      .default = municipality
    )
  ) %>%
  dplyr::arrange(municipality)

# write formatted data
write.csv(
  x = census_data_1991,
  file = "../Formatted Data/census_data_1991_formatted.csv",
  row.names = FALSE
  )

```

```{r}
#| label: census-data-1991-postwar
#| echo: FALSE

# 1991 census data assigned to post-war municipalities

census_data_1991_postwar <- bih_census_1991_postwar_modified %>%
  # drop unneeded columns
  dplyr::select(Municipality, `Population 1991`) %>%
  dplyr::mutate(
    # rename post-war municipalities to standardize names
    Municipality = dplyr::case_match(
      Municipality,
      "Domaljevac Šamac" ~ "Domaljevac-Šamac",
      "Foča (FBiH)" ~ "Foča-Ustikolina",
      "Trnovo (FBiH)" ~ "Trnovo-FBIH",
      "Pale (FBiH)" ~ "Pale-Prača",
      "Kupres (RS)" ~ "Kupres-RS",
      .default = Municipality
      )) %>%
  # manually add Bužim, Kostajnica, and Milići
  rbind(tibble::tibble(
    Municipality = c("Bužim", "Kostajnica", "Milići"),
    `Population 1991` = c(16940, 6231, 16038)
  )) %>%
  dplyr::left_join(municipality_reorg_crosswalk, by = dplyr::join_by("Municipality" == "Postwar Municipality")) %>%
  dplyr::select(
    postwar_municipality = Municipality, prewar_municipality = `Prewar Municipality`, pop_1991 = `Population 1991`, modified = Modified
  )

# create version to merge with pre-war census data
census_data_1991_new_municipalities_to_merge <- bih_census_1991_postwar_modified %>%
  dplyr::select(municipality = Municipality, population_to_subtract = `Population 1991`) %>%
  dplyr::group_by(municipality) %>%
  dplyr::summarise(population_to_subtract = sum(population_to_subtract)) %>%
  dplyr::ungroup()

# create version to append to pre-war census data
census_data_1991_new_municipalities_to_append <- bih_census_1991_postwar_modified %>%
  dplyr::select(municipality = Municipality, redist_population = `Population 1991`)

census_data_1991_postwar <- census_data_1991 %>%
  dplyr::select(municipality, population) %>%
  dplyr::left_join(census_data_1991_new_municipalities_to_merge, by = "municipality") %>%
  dplyr::mutate(
    population_to_subtract = ifelse(is.na(population_to_subtract), 0, population_to_subtract),
    population = population - population_to_subtract
    ) %>%
  dplyr::select(municipality, redist_population = population) %>%
  rbind(census_data_1991_new_municipalities_to_append) %>%
  dplyr::mutate(
    # rename post-war municipalities to standardize names
    municipality = dplyr::case_match(
      municipality,
      "Bosanska Dubica" ~ "Dubica",
      "Bosanska Gradiška" ~ "Gradiška",
      "Bosanski Brod" ~ "Brod",
      "Bosanski Novi" ~ "Novi Grad",
      "Bosanski Šamac" ~ "Šamac",
      "Gornji Vakuf" ~ "Gornji Vakuf-Uskoplje",
      "Kupres" ~ "Kupres-FBIH",
      "Prozor" ~ "Prozor-Rama",
      "Sarajevo Novo Sarajevo" ~ "Novo Sarajevo",
      "Sarajevo Vogošća" ~ "Vogošća",
      "Trnovo" ~ "Trnovo-RS",
      .default = municipality
      ))

# write formatted data
write.csv(
  x = census_data_1991_postwar,
  file = "../Formatted Data/census_data_1991_postwar_municipalities.csv",
  row.names = FALSE
  )

```

```{r}
#| label: census-data-1991-by-community
#| echo: FALSE

# 1991 census data by community (village/town/city)

census_data_1991_community <- bih_census_1991_community %>%
  dplyr::filter(
    # drop opština and country total rows
    `Mjesne Zajednice` != "TOTAL",
    # drop country total percentages row
    !is.na(`Mjesne Zajednice`)
  ) %>%
  # rename variables to English
  dplyr::select(municipality = Opština, community = `Mjesne Zajednice`, population = Ukupno,
                muslims = Muslimani, serbs = Srbi, croats = Hrvati, yugoslavs = Jugosloveni,
                other = Ostali) %>%
  dplyr::mutate(
    # calculate municipality-level percentages by ethnicity
    dplyr::across(c(muslims, serbs, croats, yugoslavs, other), ~ .x / population,
                  .names = "percent_{.col}")
  ) %>%
  dplyr::group_by(municipality) %>%
  dplyr::mutate(population_percent_of_municipality = population / sum(population)) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    # calculate polarization and fractionalization metrics
    fractionalization = 1 - rowSums(dplyr::across(dplyr::starts_with("percent_"))^2),
    polarization = 1 - 4 * rowSums(dplyr::across(dplyr::starts_with("percent_"), ~ .x * (0.5 - .x)^2))
  ) %>%
  dplyr::arrange(municipality, community)

# write formatted data
write.csv(
  x = census_data_1991_community,
  file = "../Formatted Data/census_data_1991_community_formatted.csv",
  row.names = FALSE
  )

```

### Death Data
```{r}
#| label: death-data-costalli-and-moro
#| echo: FALSE

death_data_cm <- death_data_cm_original %>%
  # rename variables
  dplyr::select(municipality = Municipality, fractionalization_cm = Fractionalization,
                polarization_cm = Polarization, ethnic_dominance_cm = `Ethnic Dominance`,
                deaths_per_population = `Deaths/Population`) %>%
  dplyr::mutate(
    # rename pre-war municipalities to standardize names
    municipality = ifelse(municipality == "Sarajevo Novo Sar.",
                          "Sarajevo Novo Sarajevo",
                          municipality)
    ) %>%
  # merge in 1991 census data to calculate absolute death values
  dplyr::full_join(census_data_1991, by = "municipality") %>%
  # calculate approximate total death counts based on 1991 census population values
  dplyr::mutate(deaths_count = population * deaths_per_population) %>%
  # drop unneeded population variables
  dplyr::select(municipality, deaths_per_population, deaths_count, fractionalization_cm,
                polarization_cm, ethnic_dominance_cm) %>%
  dplyr::arrange(municipality)

# create additional rows for new post-war municipalities
# death_data2 <- death_data_cm %>%
#   dplyr::mutate(municipality = dplyr::case_match(
#     municipality,
#     "Bosanska Krupa" ~	"Bužim",
#     "Sanski Most" ~	"Oštra Luka",
#     "Ključ" ~	"Ribnik",
#     "Bosanski Petrovac" ~	"Petrovac",
#     "Drvar" ~	"Istočni Drvar",
#     "Kupres" ~	"Kupres-RS",
#     "Jajce" ~	"Jezero",
#     "Kneževo" ~	"Dobretići",
#     "Doboj" ~	"Doboj Jug",
#     "Gračanica" ~	"Petrovo",
#     "Gradačac" ~	"Pelagićevo",
#     "Odžak" ~	"Vukosavlje",
#     "Šamac" ~	"Domaljevac-Šamac",
#     "Orašje" ~	"Donji Žabar",
#     "Lopare" ~	"Čelić",
#     "Ugljevik" ~	"Teočak",
#     "Zvornik" ~	"Sapna",
#     "Kalesija" ~	"Osmaci",
#     "Trebinje" ~	"Ravno",
#     "Stolac" ~	"Berkovići",
#     "Mostar" ~	"Istočni Mostar",
#     "Trnovo" ~	"Trnovo-RS",
#     "Foča" ~	"Foča-Ustikolina",
#     "Goražde" ~	"Novo Goražde",
#     "Pale" ~	"Pale-Prača",
#     "Ilidža" ~	"Istočna Ilidža",
#     "Sarajevo Stari Grad" ~	"Istočno Stari Grad",
#     "Sarajevo Novo Sarajevo" ~	"East New Sarajevo",
#     "Bosanski Novi" ~	"Kostajnica",
#     "Tešanj" ~	"Usora",
#     "Vlasenica" ~	"Milići",
#     .default = NA
#   )) %>%
#   dplyr::filter(!is.na(municipality))
# 
# death_data3 <- death_data_cm %>%
#   dplyr::mutate(municipality = dplyr::case_match(
#     municipality,
#     "Bosanska Krupa" ~ "Krupa na Uni",
#     "Doboj" ~ "Doboj Istok",
#     .default = NA
#   )) %>%
#   dplyr::filter(!is.na(municipality))
# 
# death_data_postwar_municipalities <- death_data_cm %>%
#   rbind(death_data2, death_data3)

# write formatted data
write.csv(
  x = death_data_cm,
  file = "../Formatted Data/deaths_per_municipality_formatted.csv",
  row.names = FALSE
  )
# write.csv(
#   x = death_data_postwar_municipalities,
#   file = "../Formatted Data/deaths_per_postwar_municipality_formatted.csv",
#   row.names = FALSE
#   )

```

```{r}
#| label: death-data-ucdp
#| echo: FALSE

# calculate total Bosnia and Herzegovina dataset
gede_bih <- gede %>%
  # restrict data to Bosnia and Herzegovina
  dplyr::filter(
    country == "Bosnia-Herzegovina"
    ) %>%
  # drop unneeded columns
  dplyr::select(-c(active_year, code_status, number_of_sources, source_article, source_office,
                   source_date, source_headline, source_original, adm_1, adm_2, geom_wkt,
                   priogrid_gid, country, country_id, region, gwnoa, gwnob))

# calculate dataset restricted to geocoded entries
gede_bih_geocoded <- gede_bih %>%
  dplyr::filter(
    # filter out location precision of 6 (country-level))
    where_prec %in% c(1, 2, 5)
    ) %>%
  # convert to an sf object
  sf::st_as_sf(
    coords = c("longitude", "latitude"),
    crs = 4326,
    remove = FALSE
    ) %>%
  # calculate the length of each event and add one so date range is inclusive of endpoints
  dplyr::mutate(
    event_length = difftime(date_end, date_start, units = "days") + 1
    )

# calculate death summary statistics
# total deaths
sum(gede_bih$best, na.rm = FALSE)

# geocoded deaths
sum(gede_bih_geocoded$best, na.rm = FALSE)

# 99.5% of deaths (best estimate) are geocoded



## prewar municipality mapping
# confirm CRSs match
sf::st_crs(gede_bih_geocoded)$proj4string
sf::st_crs(bih_prewar_municipalities_shapefile)$proj4string
# both WGS84

# spatial join
gede_municipality_prewar_assignment <- sf::st_join(
  gede_bih_geocoded,
  bih_prewar_municipalities_shapefile,
  join = st_within,
  left = TRUE
)

# collapse by municipality
gede_municipality_prewar_counts <- gede_municipality_prewar_assignment %>%
  dplyr::group_by(municipality) %>%
  dplyr::summarise(
    count_of_events = n(),
    best = sum(best, na.rm = TRUE),
    high = sum(high, na.rm = TRUE),
    low = sum(low, na.rm = TRUE)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(municipality) %>%
  # add 1991 census data
  dplyr::full_join(census_data_1991, by = "municipality") %>%
  mutate(across(c(count_of_events, best, high, low), ~replace_na(.x, 0))) %>%
  dplyr::mutate(
    deaths_per_100000 = 100000 * best / population,
    deaths_per_100000_high = 100000 * high / population,
    deaths_per_100000_low = 100000 * low / population
  ) %>%
  as.data.frame() %>%
  dplyr::select(municipality, count_of_events, best, high, low, deaths_per_100000,
                deaths_per_100000_high, deaths_per_100000_low)

# write formatted data
write.csv(
  x = gede_municipality_prewar_counts,
  file = "../Formatted Data/ucdp_deaths_prewar_municipality_formatted.csv",
  row.names = FALSE
  )



## postwar municipality mapping
bih_postwar_municipalities_shapefile <- bih_postwar_municipalities_shapefile %>%
  sf::st_make_valid()

# confirm CRSs match
sf::st_crs(gede_bih_geocoded)$proj4string
sf::st_crs(bih_postwar_municipalities_shapefile)$proj4string
# both WGS84

# spatial join
gede_municipality_postwar_assignment <- sf::st_join(
  gede_bih_geocoded,
  bih_postwar_municipalities_shapefile,
  join = st_within,
  left = TRUE
)

# collapse by municipality
gede_municipality_postwar_counts <- gede_municipality_postwar_assignment %>%
  dplyr::group_by(municipality) %>%
  dplyr::summarise(
    count_of_events = n(),
    best = sum(best, na.rm = TRUE),
    high = sum(high, na.rm = TRUE),
    low = sum(low, na.rm = TRUE)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(municipality) %>%
  as.data.frame() %>%
  dplyr::select(-geometry)

# write formatted data
write.csv(
  x = gede_municipality_postwar_counts,
  file = "../Formatted Data/ucdp_deaths_postwar_municipality_formatted.csv",
  row.names = FALSE
  )

```

## Post-War Data
This section formats data that is from after the 1992-1995 war or is based upon the post-war municipalities.

### 2013 Census Data
```{r}
#| label: census-data-2013
#| echo: FALSE

census_data_2013 <- bih_census_2013 %>%
  dplyr::rename_with(tolower) %>%
  dplyr::mutate(
    # calculate municipality-level percentages by ethnicity
    population = bosniak + croat + serb + `not declared` + other + `no answer`,
    # calculate municipality-level percentages by ethnicity
    dplyr::across(c(bosniak, serb, croat, `not declared`, other, `no answer`), ~ .x / population,
                  .names = "percent_{.col}"),
    # calculate polarization and fractionalization metrics
    fractionalization = 1 - rowSums(dplyr::across(dplyr::starts_with("percent_"))^2),
    polarization = 1 - 4 * rowSums(dplyr::across(dplyr::starts_with("percent_"), ~ .x * (0.5 - .x)^2)),
    percent_total_population = population / sum(population),
    # rename post-war municipalities to standardize names
    municipality = dplyr::case_match(
      municipality,
      "Centar Sarajevo" ~ "Sarajevo Centar",
      "Doboj - Istok" ~ "Doboj Istok",
      "Doboj - Jug" ~ "Doboj Jug",
      "Foča-FBiH" ~ "Foča-Ustikolina",
      "Foča-RS" ~ "Foča",
      "Gornji Vakuf - Uskoplje" ~ "Gornji Vakuf-Uskoplje",
      "Kneževo" ~ "Skender Vakuf",
      "Kupres-FBiH" ~ "Kupres-FBIH",
      "Kozarska Dubica" ~ "Dubica",
      "Novi Grad Sarajevo" ~ "Sarajevo Novi Grad",
      "Pale-FBiH" ~ "Pale-Prača",
      "Pale-RS" ~ "Pale",
      "Stari Grad Sarajevo" ~ "Sarajevo Stari Grad",
      "Trnovo-FBiH" ~ "Trnovo-FBIH",
      # "Brod" ~ "Bosanski Brod",
      # "Doboj - Istok" ~	"Doboj East",
      # "Doboj - Jug" ~	"Doboj South",
      # "Foča-FBiH" ~	"Foča-Ustikolina",
      # "Foča-RS" ~	"Foča",
      # "Gornji Vakuf - Uskoplje" ~	"Gornji Vakuf-Uskoplje",
      # "Gradiška" ~ "Bosanska Gradiška",
      # "Istočno Novo Sarajevo" ~	"East New Sarajevo",
      # "Kneževo" ~	"Skender Vakuf / Kneževo",
      # "Kostajnica" ~ "Bosanska Kostajnica",
      # "Kozarska Dubica" ~	"Dubica",
      # "Pale-FBiH" ~	"Pale-Prača",
      # "Pale-RS" ~	"Pale",
      .default = municipality
    )) %>%
  dplyr::rename(
    not_declared = `not declared`,
    no_answer = `no answer`,
    percent_not_declared = `percent_not declared`,
    percent_no_answer = `percent_no answer`
  ) %>%
  dplyr::arrange(municipality)

# write formatted data
write.csv(
  x = census_data_2013,
  file = "../Formatted Data/census_data_2013_formatted.csv",
  row.names = FALSE
  )

```

### Election Data
```{r}
#| label: municipal-election-results-1997
#| echo: FALSE

election_tally <- municipal_election_tally %>%
  # rename variables
  dplyr::select(municipality_code = MunCode, municipality = Municipality, party = `Party/Coalition`,
                vote_count = Votes, total_municipal_votes = TotMunVotes) %>%
  dplyr::mutate(
    # rename post-war municipalities to standardize names, including renaming all Mostar districts
    municipality = dplyr::case_match(
      municipality,
      "Bosanska Dubica / Kozarska Dubica" ~ "Dubica",
      "Bosanska Gradiška / Gradiška" ~ "Gradiška",
      "Bosanska Krupa / Krupa na Uni" ~ "Krupa na Uni",
      "Bosanski Brod / Srpski Brod" ~ "Brod",
      "Bosanski Novi / Novi Grad" ~ "Novi Grad",
      "Bosanski Petrovac / Petrovac" ~ "Petrovac",
      "Bosanski Šamac / Šamac" ~ "Šamac",
      "Bosansko Grahovo / Grahovo" ~ "Bosansko Grahovo",
      "Centar Sarajevo" ~ "Sarajevo Centar",
      "Drvar / Srpski Drvar" ~ "Istočni Drvar",
      "Foča / Srbinje" ~ "Foča",
      "Goražde / Srpsko Goražde" ~ "Novo Goražde",
      "Gornji Vakuf" ~ "Gornji Vakuf-Uskoplje",
      "Gračanica / Petrovo" ~ "Petrovo",
      "Gradačac / Pelagićevo" ~ "Pelagićevo",
      "Ilidža / Srpska Ilidža" ~ "Istočna Ilidža",
      "Jajce / Jezero" ~ "Jezero",
      "Kalesija / Osmaci" ~ "Osmaci",
      "Ključ / Ribnik" ~ "Ribnik",
      "Kupres" ~ "Kupres-FBIH",
      "Kupres / Srpski Kupres" ~ "Kupres-RS",
      "Mostar / Srpski Mostar" ~ "Istočni Mostar",
      "Novi Grad Sarajevo" ~ "Sarajevo Novi Grad",
      "Novo Sarajevo / Srpsko Novo Sarajevo" ~ "Istočno Novo Sarajevo",
      "Odžak / Vukosavlje" ~ "Vukosavlje",
      "Orašje / Srpsko Orašje" ~ "Donji Žabar",
      "Pale (RS)" ~ "Pale",
      "Prozor / Prozor-Rama" ~ "Prozor-Rama",
      "Sanski Most / Srpski Sanski Most" ~ "Oštra Luka",
      "Skender Vakuf / Kneževo" ~ "Skender Vakuf",
      "Stari Grad Sarajevo" ~ "Sarajevo Stari Grad",
      "Stari Grad Sarajevo / Srpski Stari Grad" ~ "Istočni Stari Grad",
      "Stolac / Berkovići" ~ "Berkovići",
      "Trnovo (FBIH)" ~ "Trnovo-FBIH",
      "Trnovo (RS)" ~ "Trnovo-RS",
      # "Stolac / Berkovići" ~ "Berkovići",
      # "Bosansko Grahovo / Grahovo" ~ "Bosansko Grahovo",
      # "Bosanski Brod / Srpski Brod" ~ "Bosanski Brod",
      # "Orašje / Srpsko Orašje" ~ "Donji Žabar",
      # "Foča / Srbinje" ~ "Foča",
      # "Gornji Vakuf" ~ "Gornji Vakuf-Uskoplje",
      # "Bosanska Gradiška / Gradiška" ~ "Bosanska Gradiška",
      # "Ilidža / Srpska Ilidža" ~ "Istočna Ilidža",
      # "Drvar / Srpski Drvar" ~ "Istočni Drvar",
      # "Mostar / Srpski Mostar" ~ "Istočni Mostar",
      # "Stari Grad Sarajevo / Srpski Stari Grad" ~ "Istočni Stari Grad",
      # "Novo Sarajevo / Srpsko Novo Sarajevo" ~ "East New Sarajevo",
      # "Jajce / Jezero" ~ "Jezero",
      # "Bosanska Dubica / Kozarska Dubica" ~ "Dubica",
      # "Bosanska Krupa / Krupa na Uni" ~ "Krupa na Uni",
      # "Kupres" ~ "Kupres-FBIH",
      # "Kupres / Srpski Kupres" ~ "Kupres-RS",
      # "Bosanski Novi / Novi Grad" ~ "Novi Grad",
      # "Goražde / Srpsko Goražde" ~ "Novo Goražde",
      # "Kalesija / Osmaci" ~ "Osmaci",
      # "Sanski Most / Srpski Sanski Most" ~ "Oštra Luka",
      # "Pale (RS)" ~ "Pale",
      # "Gradačac / Pelagićevo" ~ "Pelagićevo",
      # "Bosanski Petrovac / Petrovac" ~ "Petrovac",
      # "Gračanica / Petrovo" ~ "Petrovo",
      # "Prozor / Prozor-Rama" ~ "Prozor-Rama",
      # "Ključ / Ribnik" ~ "Ribnik",
      # "Bosanski Šamac / Šamac" ~ "Šamac",
      # "Trnovo (FBIH)" ~ "Trnovo-FBIH",
      # "Trnovo (RS)" ~ "Trnovo-RS",
      # "Odžak / Vukosavlje" ~ "Vukosavlje",
      # collapse portions of Mostar (FBIH) municipality
      "Mostar Jug" ~ "Mostar",
      "Mostar Jugoistok" ~ "Mostar",
      "Mostar Jugozapad" ~ "Mostar",
      "Mostar Sjever" ~ "Mostar",
      "Mostar Stari Grad" ~ "Mostar",
      "Mostar Zapad" ~ "Mostar",
      .default = municipality
    )) %>%
  # collapse results by municipality and party to summarize Mostar
  dplyr::group_by(municipality, party) %>%
  dplyr::summarise(
    municipality_code = max(municipality_code),
    vote_count = sum(vote_count),
    total_municipal_votes = sum(total_municipal_votes)
  ) %>%
  dplyr::ungroup() %>%
  # fix Mostar total municipal votes values
  dplyr::group_by(municipality) %>%
  dplyr::mutate(
    total_municipal_votes = max(total_municipal_votes)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    # create new municipality code for merged Mostar municipalities
    municipality_code = ifelse(municipality == "Mostar", "151", municipality_code),
    # calculate percentage of vote each party received per municipality
    percent_vote = vote_count / total_municipal_votes
    ) %>%
  dplyr::arrange(municipality)

# write formatted data
write.csv(
  x = election_tally,
  file = "../Formatted Data/election_tally_1997_formatted.csv",
  row.names = FALSE
  )

# post-war municipalities without data:
## Doboj Istok, Doboj Jug, Dobretići, Domaljevac-Šamac, Foča-FBiH, Kostajnica, Milići, Pale-FBiH, Ravno, Sapna, Teočak, Usora

```
    
```{r}
#| label: municipal-election-seat-allocation-1997
#| echo: FALSE

seat_tally <- municipal_seat_allocation %>%
  dplyr::rename(municipality = Municipality, canton = Canton, entity = Entity, voters = Voters,
                `SRS*` = `SRS....`
                ) %>%
  dplyr::filter(!municipality %in% c("TOTAL", "")) %>%
  dplyr::mutate(
    # rename post-war municipalities to standardize names
    municipality = dplyr::case_match(
      municipality,
      "Banovici" ~ "Banovići",
      "Bihac" ~ "Bihać",
      "Bileca" ~ "Bileća",
      "Bosanska Dubica" ~ "Dubica",
      "Bosanska Gradiska" ~ "Gradiška",
      "Bosanska Krupa RS" ~ "Krupa na Uni",
      "Bosanski Brod" ~ "Brod",
      "Bosanski Novi" ~ "Novi Grad",
      "Bosanski Petrovac RS" ~ "Petrovac",
      "Bosanski Samac" ~ "Šamac",
      "Brcko" ~ "Brčko",
      "Busovaca" ~ "Busovača",
      "Buzim" ~ "Bužim",
      "Cajnice" ~ "Čajniče",
      "Caplijna" ~ "Čapljina",
      "Celinac" ~ "Čelinac",
      "Centar Sarajevo" ~ "Sarajevo Centar",
      "Citluk" ~ "Čitluk",
      "Drvar RS" ~ "Istočni Drvar",
      "Foca" ~ "Foča",
      "Glamoc" ~ "Glamoč",
      "Gorazde" ~ "Goražde",
      "Gorazde RS" ~ "Novo Goražde",
      "Gornji Vakuf" ~ "Gornji Vakuf-Uskoplje",
      "Gracanica" ~ "Gračanica",
      "Gracanica RS" ~ "Petrovo",
      "Gradacac" ~ "Gradačac",
      "Gradacac RS" ~ "Pelagićevo",
      "Hadzici" ~ "Hadžići",
      "Ilidza" ~ "Ilidža",
      "Ilidza RS" ~ "Istočna Ilidža",
      "Ilijas" ~ "Ilijaš",
      "Jajce RS" ~ "Jezero",
      "Kalesija RS" ~ "Osmaci",
      "Kljuc" ~ "Ključ",
      "Kljuc RS" ~ "Ribnik",
      "Kotor Varos" ~ "Kotor Varoš",
      "Kresevo" ~ "Kreševo",
      "Kupres" ~ "Kupres-FBIH",
      "Kupres RS" ~ "Kupres-RS",
      "Laktasi" ~ "Laktaši",
      "Ljubuski" ~ "Ljubuški",
      "Lopare" ~ "Čelić",
      "Lopare RS" ~ "Lopare",
      "Modrica" ~ "Modriča",
      "Mostar RS" ~ "Istočni Mostar",
      "Mrkonjic Grad" ~ "Mrkonjić Grad",
      "Novi Grad Sarajevo" ~ "Sarajevo Novi Grad",
      "Novo Sarajevo RS" ~ "Istočno Novo Sarajevo",
      "Odzak" ~ "Odžak",
      "Odzak RS" ~ "Vukosavlje",
      "Orasje" ~ "Orašje",
      "Orasje RS" ~ "Donji Žabar",
      "Posusje" ~ "Posušje",
      "Prozor" ~ "Prozor-Rama",
      "Sanski Most RS" ~ "Oštra Luka",
      "Sekovici" ~ "Šekovići",
      "Sipovo" ~ "Šipovo",
      "Siroki Brijeg" ~ "Široki Brijeg",
      "Stari Grad Sarajevo" ~ "Sarajevo Stari Grad",
      "Stari Grad Sarajevo RS" ~ "Istočni Stari Grad",
      "Stolac RS" ~ "Berkovići",
      "Tesanj" ~ "Tešanj",
      "Teslic" ~ "Teslić",
      "Trnovo" ~ "Trnovo-FBIH",
      "Trnovo RS" ~ "Trnovo-RS",
      "Vares" ~ "Vareš",
      "Velika Kladusa" ~ "Velika Kladuša",
      "Visegrad" ~ "Višegrad",
      "Vogosca" ~ "Vogošća",
      "Zavidovici" ~ "Zavidovići",
      "Zepce" ~ "Žepče",
      "Zivinice" ~ "Živinice",
      # "Pale" ~ "Pale-RS",
      # "Skender Vakuf" ~ "Kneževo",
      # "Stari Grad Sarajevo RS" ~ "Istočni Stari Grad",
      # # collapse portions of Mostar (FBIH) municipality
      # "Mostar Jug" ~ "Mostar",
      # "Mostar Jugoistok" ~ "Mostar",
      # "Mostar Jugozapad" ~ "Mostar",
      # "Mostar Sjever" ~ "Mostar",
      # "Mostar Stari Grad" ~ "Mostar",
      # "Mostar Zapad" ~ "Mostar",
      .default = municipality
    )) %>%
  dplyr::arrange(municipality)

# write formatted data
write.csv(
  x = seat_tally,
  file = "../Formatted Data/seat_tally_1997_formatted.csv",
  row.names = FALSE
  )

# post-war municipalities without data:
## Doboj Istok, Doboj Jug, Dobretići, Domaljevac-Šamac, Foča-FBiH, Kostajnica, Milići, Pale-FBiH, Ravno, Sapna, Teočak, Usora

```

```{r}
#| label: nationalist-votes-allocation-1997
#| echo: FALSE

nat_votes_1997 <- nationalist_party_leanings_1997 %>%
  dplyr::rename_with(tolower) %>%
  dplyr::full_join(election_tally, by = "party") %>%
  dplyr::filter(
    !is.na(municipality),
    !municipality %in%
      c("Lista na Nivou / Razini Grada Mostara za Gradsko Vijeće",
        "Općina Mostar Jug za Gradsko Vijeće Grada Mostara",
        "Općina Mostar Jugoistok za Gradsko Vijeće Grada Mostara",
        "Općina Mostar Jugozapad za Gradsko Vijeće Grada Mostara",
        "Općina Mostar Sjever za Gradsko Vijeće Grada Mostara",
        "Općina Mostar Zapad za Gradsko Vijeće Grada Mostara")
      ) %>%
  # collapse by binary value
  dplyr::group_by(municipality, binary) %>%
  dplyr::summarise(
    # vote_count = sum(vote_count, na.rm = TRUE),
    # total_municipal_votes = max(total_municipal_votes, na.rm = TRUE),
    percent_vote = sum(percent_vote, na.rm = TRUE)
  ) %>%
  dplyr::ungroup() %>%
  tidyr::pivot_wider(names_from = binary, values_from = percent_vote) %>%
  replace(is.na(.), 0) %>%
  dplyr::arrange(municipality)

# write formatted data
write.csv(
  x = nat_votes_1997,
  file = "../Formatted Data/nationalist_votes_1997.csv",
  row.names = FALSE
  )

```

```{r}
#| label: voting-location-1997
#| echo: FALSE

voting_location <- voting_location_1997 %>%
  dplyr::select(municipality = Municipality, votes_cast_in_district = NumInDist,
                votes_cast_out_district = NumOutDist, total_votes_cast = TotalVotes) %>%
  dplyr::mutate(
    # rename post-war municipalities to standardize names, including renaming all Mostar districts
    municipality = dplyr::case_match(
      municipality,
      "Bosanska Gradiška" ~ "Gradiška",
      "Bosanski Brod" ~ "Brod",
      "Centar Sarajevo" ~ "Sarajevo Centar",
      "Doboj-Istok" ~ "Doboj Istok",
      "Doboj-Jug" ~ "Doboj Jug",
      "Foča-FBiH" ~ "Foča-Ustikolina",
      "Istočna Mostar" ~ "Istočni Mostar",
      "Kupres-FBiH" ~ "Kupres-FBIH",
      "Novi Grad Sarajevo" ~ "Sarajevo Novi Grad",
      "Pale-FBiH" ~ "Pale-Prača",
      "Pale-RS" ~ "Pale",
      "Skender Vakuf / Kneževo" ~ "Skender Vakuf",
      "Stari Grad Sarajevo" ~ "Sarajevo Stari Grad",
      "Trnovo-FBiH" ~ "Trnovo-FBIH",
      # collapse portions of Mostar (FBIH) municipality
      "Mostar Jug" ~ "Mostar",
      "Mostar Jugoistok" ~ "Mostar",
      "Mostar Jugozapad" ~ "Mostar",
      "Mostar Sjever" ~ "Mostar",
      "Mostar Stari Grad" ~ "Mostar",
      "Mostar Zapad" ~ "Mostar",
      .default = municipality
    )
  ) %>%
  dplyr::group_by(municipality) %>%
  dplyr::summarise(
    votes_cast_in_district = sum(votes_cast_in_district, na.rm = TRUE),
    votes_cast_out_district = sum(votes_cast_out_district, na.rm = TRUE),
    total_votes_cast = sum(total_votes_cast, na.rm = TRUE),
    perc_cast_in = votes_cast_in_district / total_votes_cast,
    perc_cast_out = votes_cast_out_district / total_votes_cast,
  ) %>%
  dplyr::ungroup() %>%
  dplyr::arrange(municipality)

# write formatted data
write.csv(
  x = voting_location,
  file = "../Formatted Data/voting_location_1997.csv",
  row.names = FALSE
  )

```

## Distances from Neighboring Countries
```{r}
#| label: municipal-distances-from-neighbors
#| echo: FALSE

# create shapefile of united Serbia and Montenegro (rump Yugoslavia)
yug_shapefile <- srb_shapefile %>%
  rbind(mne_shapefile) %>%
  rbind(xko_shapefile) %>%
  sf::st_as_sf() %>%
  sf::st_union() %>%
  sf::st_as_sf()

## pre-war data
mun_distances_prewar <- bih_prewar_municipalities_shapefile %>%
  dplyr::mutate(
    # calculate distances from each municipality to HRV, SRB, MNE and rump YUG, as well as the
    # inter-ethnic boundary line
    dist_hrv = units::set_units(sf::st_distance(geom, sf::st_as_sf(hrv_shapefile))[,1], km),
    dist_srb = units::set_units(sf::st_distance(geom, sf::st_as_sf(srb_shapefile))[,1], km),
    dist_mne = units::set_units(sf::st_distance(geom, sf::st_as_sf(mne_shapefile))[,1], km),
    dist_yug = units::set_units(sf::st_distance(geom, sf::st_as_sf(yug_shapefile))[,1], km),
    dist_internal_border = units::set_units(sf::st_distance(geom, bih_postwar_iebl_shapefile)[,1], km)
  ) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(dist_other_country = min(dist_hrv, dist_yug)) %>%
  as.data.frame() %>%
  dplyr::select(municipality, dist_hrv, dist_srb, dist_mne, dist_yug, dist_other_country,
                dist_internal_border)

# write formatted data
write.csv(
  x = mun_distances_prewar,
  file = "../Formatted Data/prewar_municipality_distances.csv",
          row.names = FALSE)

## post-war data
mun_distances_postwar <- bih_postwar_municipalities_shapefile %>%
  dplyr::mutate(
    # calculate distances from each municipality to HRV, SRB, MNE and rump YUG, as well as the
    # inter-ethnic boundary line
    dist_hrv = units::set_units(sf::st_distance(geom, sf::st_as_sf(hrv_shapefile))[,1], km),
    dist_srb = units::set_units(sf::st_distance(geom, sf::st_as_sf(srb_shapefile))[,1], km),
    dist_mne = units::set_units(sf::st_distance(geom, sf::st_as_sf(mne_shapefile))[,1], km),
    dist_yug = units::set_units(sf::st_distance(geom, sf::st_as_sf(yug_shapefile))[,1], km),
    dist_internal_border = units::set_units(sf::st_distance(geom, bih_postwar_iebl_shapefile)[,1], km)
  ) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(dist_other_country = min(dist_hrv, dist_yug)) %>%
  as.data.frame() %>%
  dplyr::select(municipality, dist_hrv, dist_srb, dist_mne, dist_yug, dist_other_country,
                dist_internal_border)

# write formatted data
write.csv(
  x = mun_distances_postwar,
  file = "../Formatted Data/postwar_municipality_distances.csv",
  row.names = FALSE
  )

```

# Mines and Minefields
```{r}
#| label: canton-minefields-1997
#| echo: FALSE

# Calculate estimates for minefield prevalence at the cantonal level in 1997.

## Calculate Canton Areas
## Calculate areas of the 10 cantons in the Federation, Brčko, and Republika Srpska.
canton_area <- bih_postwar_cantons_shapefile %>%
  # for cantons in Republika Srpska, collapse into "Republika Srpska"
  dplyr::mutate(
    canton = dplyr::case_when(
      entity == "Republika Srpska" ~ "Republika Srpska",
      .default = canton
    )) %>%
  dplyr::group_by(canton) %>%
  dplyr::summarise(.groups = 'drop') %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    canton_area = units::set_units(sf::st_area(geom), km^2)
  )

## Format Minefield Cantonal Data
landmines1 <- landmines %>%
  dplyr::select(canton = `Federation:`, minefields = Minefields) %>%
  replace(is.na(.), 0) %>%
  dplyr::mutate(canton = ifelse(canton == "Republika Srpska:", "Republika Srpska", canton)) %>%
  dplyr::filter(canton != "Total:") %>%
  # remove extra whitespace on either end of any strings and replace internal whitespace with
  # standard space character
  dplyr::mutate(
    canton = stringr::str_squish(canton)
  ) %>%
  dplyr::full_join(canton_area, by = "canton") %>%
  dplyr::mutate(
    minefields_per_sq_km = minefields / canton_area
  )

## Replace Republika Srpska estimates with estimates for individual cantons.

# pull RS minefield density estimate
rs_minefield_density_est <- landmines1$minefields_per_sq_km[landmines1$canton == "Republika Srpska"]

landmines_rs_cantons <- data.frame(
  canton = c("Banja Luka", "Bijeljina", "Doboj", "Foča", "Istočno Sarajevo",
             "Trebinje", "Vlasenica"),
  minefields_per_sq_km = rep(rs_minefield_density_est, 7)
)

# append data
landmines1 <- landmines1 %>%
  dplyr::select(canton, minefields_per_sq_km) %>%
  dplyr::filter(!canton %in% c("Republika Srpska", "Brčko")) %>%
  rbind(landmines_rs_cantons)

# estimate Brčko
# 59.6 sq. km in Brčko are suspected areas - 12.04% of territory.
# 6.36% estimate in FBiH and 1.71% in Republika Srpska.

# Bosnia-wide estimate of 3 million landmines; 152 mines per square mile.

# Divide 3 million landmines from number of minefields in data to get estimate of mines per
# minefield.
total_minefields <- sum(landmines$Minefields, na.rm = TRUE)
average_mines_per_minefield = 3000000 / total_minefields

# estimated 152 mines per square mile
# convert to sq km
mines_per_sq_km_est <- 152 / 2.59

# pull Brčko size in sq km
brcko_area <- canton_area$canton_area[canton_area$canton == "Brčko"]

brcko_mine_estimate <- mines_per_sq_km_est * brcko_area
brcko_minefield_estimate <- brcko_mine_estimate / average_mines_per_minefield
brcko_minefield_density_estimate <- units::set_units(as.numeric(brcko_minefield_estimate / brcko_area), 1/km^2)

# add Brčko estimate
landmines1 <- landmines1 %>%
  rbind(data.frame(
    canton = "Brčko",
    minefields_per_sq_km = brcko_minefield_density_estimate
  )) %>%
  dplyr::arrange(canton)

# write formatted data
write.csv(
  x = landmines1,
  file = "../Formatted Data/landmines_1997.csv",
  row.names = FALSE
  )

```

```{r}
#| label: municipal-minefield-coverage-2004
#| echo: FALSE

# Using a 2004 map showing approximate minefield areas, cleared minefield areas, post-war mine
# explosions, and frontlines of the battlefield, use the active and cleared minefield areas and
# post-war mine explosion location data to approximate minefield areas. Use this geocoded data
# to approximate proportion of post-war municipal areas at risk for landmine infection.

bih_postwar_municipalities_shapefile <- bih_postwar_municipalities_shapefile %>%
  sf::st_make_valid()

minefield_2004_shapefile <- minefield_2004_shapefile %>%
  sf::st_make_valid()

ggplot() +
  # Bottom layer (Base map)
  geom_sf(data = bih_postwar_municipalities_shapefile, fill = "gray80", color = "white") + 
  # Top layer (Overlay with transparency)
  geom_sf(data = minefield_2004_shapefile, fill = "blue", alpha = 0.5, color = "black") + 
  theme_minimal()

# confirm CRSs match
sf::st_crs(minefield_2004_shapefile)$proj4string
sf::st_crs(bih_postwar_municipalities_shapefile)$proj4string
# both WGS84

bih_minefields_2004 <- sf::st_intersection(
  minefield_2004_shapefile,
  bih_postwar_municipalities_shapefile
  )

# collapse by municipality
bih_minefields_2004_perc_area <- bih_minefields_2004 %>%
  sf::st_make_valid() %>%
  dplyr::mutate(
    # calculate updated geometry area
    area = units::set_units(sf::st_area(geom), km^2)
    ) %>%
  dplyr::group_by(municipality) %>%
  dplyr::summarise(
    minefield_area = sum(area, na.rm = TRUE),
    municipality_area = min(mun_area, na.rm = TRUE)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    minefield_perc = units::drop_units(minefield_area / municipality_area),
  ) %>%
  dplyr::arrange(municipality) %>%
  as.data.frame() %>%
  dplyr::full_join(
    bih_postwar_municipalities_shapefile,
    by = "municipality"
    ) %>%
  dplyr::select(municipality, minefield_area, minefield_perc) %>%
  replace(is.na(.), 0)

# write formatted data
write.csv(
  x = bih_minefields_2004_perc_area,
  file = "../Formatted Data/landmines_2004.csv",
  row.names = FALSE
  )

```

```{r}
#| label: canton-minefield-coverage-2004
#| echo: FALSE

bih_postwar_cantons_shapefile <- bih_postwar_cantons_shapefile %>%
  dplyr::mutate(
    canton_area = units::set_units(sf::st_area(geom), km^2)
  )

bih_minefields_2004_canton <- sf::st_intersection(
  minefield_2004_shapefile,
  bih_postwar_cantons_shapefile
  )

# collapse by canton
bih_minefields_2004_perc_canton_area <- bih_minefields_2004_canton %>%
  sf::st_make_valid() %>%
  dplyr::mutate(
    # calculate updated geometry area
    minefield_area = units::set_units(sf::st_area(geom), km^2)
  ) %>%
  dplyr::group_by(canton) %>%
  dplyr::summarise(
    minefield_area = sum(minefield_area, na.rm = TRUE),
    canton_area = min(canton_area, na.rm = TRUE)
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(
    minefield_perc = units::drop_units(minefield_area / canton_area),
  ) %>%
  dplyr::arrange(canton) %>%
  as.data.frame() %>%
  dplyr::full_join(
    bih_postwar_cantons_shapefile,
    by = "canton"
  ) %>%
  dplyr::select(canton, minefield_area, minefield_perc) %>%
  replace(is.na(.), 0) %>%
  dplyr::mutate(
    canton = stringr::str_squish(canton)
  )

# write formatted data
write.csv(
  x = bih_minefields_2004_perc_canton_area,
  file = "../Formatted Data/landmines_2004_canton.csv",
  row.names = FALSE
  )

```
