---
title: "03-Merge_Data"
format: html
---

This script loads all data formatted in 02-Format_Data.qmd and compiles it into three datasets:
(1) Pre-War data - data compiled by pre-war municipalities
(2) Post-War data - data compiled by post-war municipalities
(3) Pre-War data assigned to post-war municipalities - data compiled at a pre-war municipal level assigned to post-war municipalities; this version will have the same data for municipalities that were originally one municipality and sums will not add up to the full value due to double counting

## Loading
```{r}
#| label: load-libraries
#| echo: FALSE

library(sf)
library(jsonlite)
library(readxl)
library(openxlsx)
library(dplyr)

```

```{r}
#| label: load-data
#| echo: FALSE

### PRE-WAR DATA ###

# shapefile
bih_prewar_municipalities_shapefile <- sf::read_sf(
  "../Shape Files/Formatted Shape Files/bih_prewar_municipalities_shapefile_formatted.gpkg",
  quiet = TRUE
  )

# 1991 census data
census_data_1991 <- read.csv("../Formatted Data/census_data_1991_formatted.csv")

# death data
death_data_cm <- read.csv("../Formatted Data/deaths_per_municipality_formatted.csv")

# municipality distances
mun_distances_prewar <- read.csv("../Formatted Data/prewar_municipality_distances.csv")

# pre-war x post-war municipality crosswalk
municipality_crosswalk <- readxl::read_xlsx("../Data/bih_municipality_reorg_crosswalk.xlsx")

# gede deaths pre-war municipalities
gede_municipality_prewar_counts <- read.csv("../Formatted Data/ucdp_deaths_prewar_municipality_formatted.csv")

### POST-WAR DATA ###

# shapefile
bih_postwar_municipalities_shapefile <- sf::read_sf(
  "../Shape Files/Formatted Shape Files/bih_postwar_municipalities_shapefile_formatted.gpkg",
  quiet = TRUE
)

# 2013 census data
census_data_2013 <- read.csv("../Formatted Data/census_data_2013_formatted.csv")

# 1991 census data applied to post-war municipalities
census_data_1991_redist <- read.csv("../Formatted Data/census_data_1991_postwar_municipalities.csv")

# nationalist/non-nationalist 1997 vote proportions
nationalist_votes <- read.csv("../Formatted Data/nationalist_votes_1997.csv")

# 1997 voting location
voting_location <- read.csv("../Formatted Data/voting_location_1997.csv")

# municipality distances
mun_distances_postwar <- read.csv("../Formatted Data/postwar_municipality_distances.csv")

# post-war municipality metadata
postwar_metadata <- readxl::read_xlsx("../Data/bih_postwar_municipality_metadata.xlsx")

# minefields 2004
bih_minefields_2004_perc_area <- read.csv("../Formatted Data/landmines_2004.csv")

gede_municipality_postwar_counts <- read.csv("../Formatted Data/ucdp_deaths_postwar_municipality_formatted.csv")

### POST-WAR CANTON DATA ###

# shapefile
bih_postwar_cantons_shapefile <- sf::read_sf(
  "../Shape Files/Formatted Shape Files/bih_postwar_cantons_shapefile_formatted.gpkg",
  quiet = TRUE
)

# minefields 1997
landmines1 <- read.csv("../Formatted Data/landmines_1997.csv")

# minefields 2004
bih_minefields_2004_perc_canton_area <- read.csv("../Formatted Data/landmines_2004_canton.csv")

```

## Pre-War Data
```{r}
#| label: prewar-data
#| echo: FALSE

prewar_data_shapefile <- bih_prewar_municipalities_shapefile %>%
  # merge in 1991 census data
  dplyr::full_join(census_data_1991, by = "municipality") %>%
  # merge in death data from c&m
  dplyr::full_join(death_data_cm, by = "municipality") %>%
  # merge in death data from udcp
  dplyr::full_join(gede_municipality_prewar_counts, by = "municipality") %>%
  # merge in municipal distances from neighboring countries / iebl
  dplyr::full_join(mun_distances_prewar, by = "municipality") %>%
  # select relevant variables and rename to be publication-ready names
  dplyr::select(
    Municipality = municipality,
    `Municipality Area, Pre-War` = mun_area,
    `Municipality Perimeter, Pre-War` = mun_perimeter,
    `Population, 1991` = population,
    `Muslim Population, 1991` = muslims,
    `Croat Population, 1991` = croats,
    `Serb Population, 1991` = serbs,
    `Yugoslav Population, 1991` = yugoslavs,
    `Other Population, 1991` = other,
    `Muslim Population Percentage, 1991` = percent_muslims,
    `Croat Population Percentage, 1991` = percent_croats,
    `Serb Population Percentage, 1991` = percent_serbs,
    `Yugoslav Population Percentage, 1991` = percent_yugoslavs,
    `Other Population Percentage, 1991` = percent_other,
    `Ethnic Fractionalization, 1991` = fractionalization,
    `Ethnic Polarization, 1991` = polarization,
    `Percent of Population, 1991` = percent_total_population,
    `Deaths per Population` = deaths_per_population,
    `Number of Deaths` = deaths_count,
    `Count of Events` = count_of_events,
    `Number of Deaths, best estimate` = best,
    `Number of Deaths, high estimate` = high,
    `Number of Deaths, low estimate` = low,
    `Deaths per 100,000, UCDP` = deaths_per_100000,
    `Deaths per 100,000, high, UCDP` = deaths_per_100000_high,
    `Deaths per 100,000, low, UCDP` = deaths_per_100000_low,
    `Ethnic Dominance` = ethnic_dominance_cm,
    `Distance to Croatia, Pre-War` = dist_hrv,
    `Distance to Serbia, Pre-War` = dist_srb,
    `Distance to Montenegro, Pre-War` = dist_mne,
    `Distance to Yugoslavia, Pre-War` = dist_yug,
    `Distance to Another Country, Pre-War` = dist_other_country,
    `Distance to IEBL, Pre-War` = dist_internal_border
  ) %>%
  # arrange by municipality name
  dplyr::arrange(Municipality)

# create df version of the data
prewar_data <- prewar_data_shapefile %>%
  as.data.frame() %>%
  dplyr::select(-geom)

# write shapefile data
sf::write_sf(
  obj = prewar_data_shapefile,
  dsn = "../Shape Files/Formatted Shape Files/prewar_data.gpkg"
  )

# write formatted data
openxlsx::write.xlsx(
  x = prewar_data,
  file = "../Formatted Data/data_prewar.xlsx"
)

```

# Post-War Data
```{r}
#| label: postwar-data
#| echo: FALSE

postwar_data_shapefile <- bih_postwar_municipalities_shapefile %>%
  # drop unneeded variables
  dplyr::select(-c(canton, entity, sarajevo_district, split_by_iebl)) %>%
  # merge in post-war municipality metadata
  dplyr::full_join(postwar_metadata, by = "municipality") %>%
  # merge in 2013 census data
  dplyr::full_join(census_data_2013, by = "municipality") %>%
  # merge in 1991 census data, by 2013 municipality
  dplyr::full_join(census_data_1991_redist, by = "municipality") %>%
  # merge in received nationalist voting data
  dplyr::full_join(nationalist_votes, by = "municipality") %>%
  # merge in voting location data
  dplyr::full_join(voting_location, by = "municipality") %>%
  # merge in municipal distances from neighboring countries / iebl
  dplyr::full_join(mun_distances_postwar, by = "municipality") %>%
  # merge in landmine 2004 data
  dplyr::left_join(landmines2, by = "municipality") %>%
  # select relevant variables and rename to be publication-ready names
  dplyr::select(
    Municipality = municipality,
    Canton = canton,
    Entity = entity,
    `Sarajevo District` = `sarajevo district`,
    `Municipal Borders Modified` = `split by line`,
    `Municipality Area, Post-War` = mun_area,
    `Municipality Perimeter, Post-War` = mun_perimeter,
    `Population, 1991 Redistributed` = redist_population,
    `Population, 2013` = population,
    `Bosniak Population, 2013` = bosniak,
    `Croat Population, 2013` = croat,
    `Serb Population, 2013` = serb,
    `Not Declared Population, 2013` = not_declared,
    `Other Population, 2013` = other,
    `No Answer Population, 2013` = no_answer,
    `Bosniak Population Percentage, 2013` = percent_bosniak,
    `Croat Population Percentage, 2013` = percent_croat,
    `Serb Population Percentage, 2013` = percent_serb,
    `Not Declared Population Percentage, 2013` = percent_not_declared,
    `Other Population Percentage, 2013` = percent_other,
    `No Answer Population Percentage, 2013` = percent_no_answer,
    `Ethnic Fractionalization, 2013` = fractionalization,
    `Ethnic Polarization, 2013` = polarization,
    `Percent of Population, 2013` = percent_total_population,
    `Proportion of Votes for Nationalist Parties, 1997` = Nationalist,
    `Proportion of Votes for Non-Nationalist Parties, 1997` = Non.Nationalist,
    `Proportion of Votes for Independent Parties, 1997` = Independent,
    `Proportion of Votes for Other Parties, 1997` = Other,
    `Proportion of Invalid Ballots, 1997` = Invalid.Ballot,
    `Number of Votes Cast In-District, 1997` = votes_cast_in_district,
    `Number of Votes Cast Out-District, 1997` = votes_cast_out_district,
    `Total Votes Cast, 1997` = total_votes_cast,
    `Proportion of Votes Cast In-District, 1997` = perc_cast_in,
    `Proportion of Votes Cast Out-District, 1997` = perc_cast_out,
    `Distance to Croatia, Post-War` = dist_hrv,
    `Distance to Serbia, Post-War` = dist_srb,
    `Distance to Montenegro, Post-War` = dist_mne,
    `Distance to Yugoslavia, Post-War` = dist_yug,
    `Distance to Another Country, Post-War` = dist_other_country,
    `Distance to IEBL, Post-War` = dist_internal_border,
    `Minefield Area` = minefield_area,
    `Minefield Percentage of Municipalty` = minefield_perc
  ) %>%
  # code binaries for entities - Brčko is coded as belonging to both
  dplyr::mutate(
    `Republika Srpska Municipality` = ifelse(Entity %in% c("Republika Srpska", "Brčko"), 1, 0),
    `Federation Municipality` = ifelse(Entity != "Republika Srpska", 1, 0),
  ) %>%
  dplyr::arrange(Municipality)

# create df version of the data
postwar_data <- postwar_data_shapefile %>%
  as.data.frame() %>%
  dplyr::select(-geom)

# write shapefile data
sf::write_sf(
  obj = postwar_data_shapefile,
  dsn = "../Shape Files/Formatted Shape Files/postwar_data.gpkg"
  )

# write formatted data
openxlsx::write.xlsx(
  x = postwar_data,
  file = "../Formatted Data/data_postwar.xlsx"
)

```

## Post-War Canton Data
```{r}
#| label: postwar-canton-data
#| echo: FALSE

postwar_canton_data_shapefile <- bih_postwar_cantons_shapefile %>%
  # remove extra whitespace on either end of any strings and replace internal whitespace with
  # standard space character
  dplyr::mutate(
    canton = stringr::str_squish(canton),
    entity = stringr::str_squish(entity)
  ) %>%
  # merge in landmine 1997 data
  dplyr::full_join(landmines1, by = "canton") %>%
  # merge in landmine 2004 data
  dplyr::full_join(bih_minefields_2004_perc_canton_area, by = "canton") %>%
  # select relevant variables and rename to be publication-ready names
  dplyr::select(
    Canton = canton,
    Entity = entity,
    `Canton Area, Post-War` = canton_area,
    `Canton Perimeter, Post-War` = canton_perimeter,
    `Minefields per Sq. km, Canton` = minefields_per_sq_km,
    `Minefield Area, Canton` = minefield_area,
    `Minefield Percentage of Canton` = minefield_perc
  ) %>%
  dplyr::arrange(Canton)

# create df version of the data
postwar_canton_data <- postwar_canton_data_shapefile %>%
  as.data.frame() %>%
  dplyr::select(-geom)

# write shapefile data
sf::write_sf(
  obj = postwar_canton_data_shapefile,
  dsn = "../Shape Files/Formatted Shape Files/postwar_canton_data.gpkg"
  )

# write formatted data
openxlsx::write.xlsx(
  x = postwar_canton_data,
  file = "../Formatted Data/data_postwar_canton.xlsx"
)

```

# Pre-War Data Assigned to Post-War Municipalities
```{r}
#| label: prewar-data-postwar-municipalities
#| echo: FALSE

prewar_data_assigned_to_postwar_shapefile <- bih_postwar_municipalities_shapefile %>%
  # as.data.frame() %>%
  dplyr::select(Municipality = municipality) %>%
  # merge in pre- and post-war municipality crosswalk
  dplyr::full_join(municipality_crosswalk,
                   dplyr::join_by("Municipality" == "Postwar Municipality")) %>%
  # merge in pre-war data
  dplyr::full_join(prewar_data, dplyr::join_by("Prewar Municipality" == "Municipality"))
  
# create df version of the data
prewar_data_assigned_to_postwar <- prewar_data_assigned_to_postwar_shapefile %>%
  as.data.frame() %>%
  dplyr::select(-geom)

# write shapefile data
sf::write_sf(
  obj = prewar_data_assigned_to_postwar,
  dsn = "../Shape Files/Formatted Shape Files/prewar_assigned_to_postwar_data.gpkg"
  )

# write formatted data
openxlsx::write.xlsx(
  prewar_data_assigned_to_postwar,
  "../Formatted Data/data_prewar_assigned_postwar.xlsx"
  )

```
