---
title: "04-GLMs"
format: html
---

...

## Loading
```{r}
#| label: load-libraries
#| echo: FALSE

library(readxl)
library(modelsummary)
library(tidyverse)

```

```{r}
#| label: load-data
#| echo: FALSE

# pre-war data
prewar_data <- readxl::read_xlsx("../Formatted Data/data_prewar.xlsx")

# post-war data
postwar_data <- readxl::read_xlsx("../Formatted Data/data_postwar.xlsx")

# post-war canton data
postwar_data_canton <- readxl::read_xlsx("../Formatted Data/data_postwar_canton.xlsx")

# pre-war data assigned to post-war municipalities
prewar_data_assigned_to_postwar <- readxl::read_xlsx(
  "../Formatted Data/data_prewar_assigned_postwar.xlsx"
  )

```

```{r}
#| label: format-data
#| echo: FALSE

df <- postwar_data %>%
  dplyr::full_join(prewar_data_assigned_to_postwar, by = "Municipality") %>%
  dplyr::full_join(postwar_data_canton, by = "Canton") %>%
  dplyr::filter(!is.na(`Proportion of Votes for Non-Nationalist Parties, 1997`)) %>%
  # calculate additional variables
  dplyr::mutate(
    `Population Density, 1991` = `Population, 1991` / `Municipality Area, Pre-War`,
    `Population Density, 1991 Redistributed` = `Population, 1991 Redistributed` / `Municipality Area, Post-War`,
    `Population Density, 1991 Redistributed Logged` = log(`Population Density, 1991 Redistributed`),
    `Population, 1991 Logged` = log(`Population, 1991`),
    `Population, 1991 Redistributed Logged` = log(`Population, 1991 Redistributed`),
    `Population, 2013 Logged` = log(`Population, 2013`),
    `Change in Population` = (`Population, 2013` - `Population, 1991 Redistributed`) / `Population, 1991 Redistributed`,
    `Change in Ethnic Fractionalization, 1991 to 2013` = `Ethnic Fractionalization, 2013` -
      `Ethnic Fractionalization, 1991`,
    `Change in Ethnic Polarization, 1991 to 2013` = `Ethnic Polarization, 2013` -
      `Ethnic Polarization, 1991`,
    `Change in % Bosniaks, 1991 to 2013` = `Bosniak Population Percentage, 2013` -
      `Muslim Population Percentage, 1991`,
    `Change in % Croats, 1991 to 2013` = `Croat Population Percentage, 2013` -
      `Croat Population Percentage, 1991`,
    `Change in % Serbs, 1991 to 2013` = `Serb Population Percentage, 2013` -
      `Serb Population Percentage, 1991`,
    `Absolute Change in Bosniaks, Croats, and Serbs, 1991 to 2013` = abs(`Change in % Bosniaks, 1991 to 2013`) +
      abs(`Change in % Croats, 1991 to 2013`) + abs(`Change in % Serbs, 1991 to 2013`),
    `Turnout Rate, 1997` = `Total Votes Cast, 1997` / `Population, 1991 Redistributed`,
    `Estimated Number of Minefields` = `Minefields per Sq. km, Canton` * `Municipality Area, Post-War`,
    `Municipality Area, Pre-War Logged` = log(`Municipality Area, Pre-War`),
    `Municipality Area, Post-War Logged` = log(`Municipality Area, Post-War`)
    )

# pull Republika Srpska and Federation exclusive datasets, with Brčko included in both

df_rs <- df %>%
  dplyr::filter(`Republika Srpska Municipality` == 1)

df_fbih <- df %>%
  dplyr::filter(`Federation Municipality` == 1)

```

```{r}
#| label: correlations
#| echo: FALSE

# correlation matrix
cor_matrix <- cor(df %>%
                    dplyr::select(
      `Deaths per Population`, `Minefields per Sq. km, Canton`, `Ethnic Fractionalization, 1991`,
      `Change in Ethnic Fractionalization, 1991 to 2013`,
      `Proportion of Votes Cast Out-District, 1997`, `Turnout Rate, 1997`,
      `Yugoslav Population Percentage, 1991`, `Other Population Percentage, 1991`,
      `Absolute Change in Bosniaks, Croats, and Serbs, 1991 to 2013`,
      `Distance to Croatia, Post-War`, `Distance to Yugoslavia, Post-War`,
      `Distance to IEBL, Post-War`, `Population Density, 1991 Redistributed Logged`,
      `Population, 1991 Redistributed Logged`, `Change in Population`, `Federation Municipality`,
      `Sarajevo District`, `Muslim Population Percentage, 1991`,
      `Change in % Bosniaks, 1991 to 2013`, `Croat Population Percentage, 1991`,
      `Change in % Croats, 1991 to 2013`, `Serb Population Percentage, 1991`,
      `Change in % Serbs, 1991 to 2013`
        ))
  
# NOTE: var1-var2 and var2-var1 will both be listed
cor_matrix_long <- cor_matrix %>%
  as.data.frame() %>%
  tibble::rownames_to_column("var1") %>%
  tidyr::pivot_longer(2:(ncol(cor_matrix) + 1), names_to = "var2", values_to = "correlation") %>%
  # drop rows where var1 == var2 (diagonal in the correlation matrix)
  dplyr::filter(var1 != var2) %>%
  # create absolute correlation, so both high and low correlation extremes can be seen together
  dplyr::mutate(cor_abs = abs(correlation))

```

## Non-Nationalist Vote Share GLMs
* A1 (a1) - Country-wide / Bosniak / Fractionalization
* A1-RS (a1r) - Republika Srpska / Bosniak / Fractionalization
* A1-FBiH (a1f) - Federation / Bosniak / Fractionalization

* A2 (a2) - Country-wide / Croat / Fractionalization
* A2-RS (a2r) - Republika Srpska / Croat / Fractionalization
* A2-FBiH (a2f) - Federation / Croat / Fractionalization

* A3 (a3) - Country-wide / Serb / Fractionalization
* A3-RS (a3r) - Republika Srpska / Serb / Fractionalization
* A3-FBiH (a3f) - Federation / Serb / Fractionalization

* A4 (a4) - Country-wide / Bosniak / Polarization
* A4-RS (a4r) - Republika Srpska / Bosniak / Polarization
* A4-FBiH (a4f) - Federation / Bosniak / Polarization

* A5 (a5) - Country-wide / Croat / Polarization
* A5-RS (a5r) - Republika Srpska / Croat / Polarization
* A5-FBiH (a5f) - Federation / Croat / Polarization

* A6 (a6) - Country-wide / Serb / Polarization
* A6-RS (a6r) - Republika Srpska / Serb / Polarization
* A6-FBiH (a6f) - Federation / Serb / Polarization

### A1 Models
```{r}
#| label: a1
#| echo: FALSE

# all municipalities
model_a1 <- glm(`Proportion of Votes for Non-Nationalist Parties, 1997` ~
                  `Deaths per Population` +
                  # `Minefields per Sq. km, Canton` +
                  `Minefield Percentage of Municipalty` +
                  `Ethnic Fractionalization, 1991` +
                  `Change in Ethnic Fractionalization, 1991 to 2013` +
                  `Proportion of Votes Cast Out-District, 1997` + `Turnout Rate, 1997` +
                  `Yugoslav Population Percentage, 1991` + `Other Population Percentage, 1991` +
                  `Absolute Change in Bosniaks, Croats, and Serbs, 1991 to 2013` +
                  `Distance to Croatia, Post-War` + `Distance to Yugoslavia, Post-War` +
                  `Distance to IEBL, Post-War` + `Population Density, 1991 Redistributed Logged` +
                  `Population, 1991 Redistributed Logged` + `Change in Population` +
                  `Federation Municipality` + `Sarajevo District` + 
                  `Muslim Population Percentage, 1991` + `Change in % Bosniaks, 1991 to 2013`,
                data = df, family = gaussian()
)
summary(model_a1)
car::vif(model_a1)

# Republika Srpska
model_a1r <- glm(`Proportion of Votes for Non-Nationalist Parties, 1997` ~
                   `Deaths per Population` +
                   `Minefield Percentage of Municipalty` +
                   `Ethnic Fractionalization, 1991` +
                   `Change in Ethnic Fractionalization, 1991 to 2013` +
                   `Proportion of Votes Cast Out-District, 1997` + `Turnout Rate, 1997` +
                   `Yugoslav Population Percentage, 1991` + `Other Population Percentage, 1991` +
                   `Absolute Change in Bosniaks, Croats, and Serbs, 1991 to 2013` +
                   `Distance to Croatia, Post-War` + `Distance to Yugoslavia, Post-War` +
                   `Distance to IEBL, Post-War` + `Population, 1991 Redistributed Logged` +
                   `Change in Population` + `Municipality Area, Post-War Logged` +
                   `Sarajevo District` + `Muslim Population Percentage, 1991` +
                   `Change in % Bosniaks, 1991 to 2013`,
                 data = df_rs, family = gaussian())
summary(model_a1r)
car::vif(model_a1r)

# Federation
model_a1f <- glm(`Proportion of Votes for Non-Nationalist Parties, 1997` ~
                   `Deaths per Population` +
                   # `Minefields per Sq. km, Canton` +
                   `Minefield Percentage of Municipalty` +
                   `Ethnic Fractionalization, 1991` +
                   `Change in Ethnic Fractionalization, 1991 to 2013` +
                   `Proportion of Votes Cast Out-District, 1997` + `Turnout Rate, 1997` +
                   `Yugoslav Population Percentage, 1991` + `Other Population Percentage, 1991` +
                   `Absolute Change in Bosniaks, Croats, and Serbs, 1991 to 2013` +
                   `Distance to Croatia, Post-War` + `Distance to Yugoslavia, Post-War` +
                   `Distance to IEBL, Post-War` + `Population, 1991 Redistributed Logged` +
                   `Change in Population` + `Municipality Area, Post-War Logged` +
                   `Sarajevo District` + `Muslim Population Percentage, 1991` +
                   `Change in % Bosniaks, 1991 to 2013`,
                 data = df_fbih, family = gaussian())
summary(model_a1f)
car::vif(model_a1f)

```

### A2 Models
```{r}
#| label: a2
#| echo: FALSE

# all municipalities
model_a2 <- glm(`Proportion of Votes for Non-Nationalist Parties, 1997` ~
                  `Deaths per Population` + `Minefields per Sq. km, Canton` + `Ethnic Fractionalization, 1991` +
                  `Change in Ethnic Fractionalization, 1991 to 2013` +
                  `Proportion of Votes Cast Out-District, 1997` + `Turnout Rate, 1997` +
                  `Yugoslav Population Percentage, 1991` + `Other Population Percentage, 1991` +
                  `Absolute Change in Bosniaks, Croats, and Serbs, 1991 to 2013` +
                  `Distance to Croatia, Post-War` + `Distance to Yugoslavia, Post-War` +
                  `Distance to IEBL, Post-War` + `Population, 1991 Redistributed Logged` +
                  `Change in Population` + `Municipality Area, Post-War Logged` +
                  `Federation Municipality` + `Sarajevo District` +
                  `Croat Population Percentage, 1991` + `Change in % Croats, 1991 to 2013`,
                data = df, family = gaussian())
summary(model_a2)
car::vif(model_a2)

# Republika Srpska
model_a2r <- glm(`Proportion of Votes for Non-Nationalist Parties, 1997` ~
                   `Deaths per Population` + `Ethnic Fractionalization, 1991` +
                   `Change in Ethnic Fractionalization, 1991 to 2013` +
                   `Proportion of Votes Cast Out-District, 1997` + `Turnout Rate, 1997` +
                   `Yugoslav Population Percentage, 1991` + `Other Population Percentage, 1991` +
                   `Absolute Change in Bosniaks, Croats, and Serbs, 1991 to 2013` +
                   `Distance to Croatia, Post-War` + `Distance to Yugoslavia, Post-War` +
                   `Distance to IEBL, Post-War` + `Population, 1991 Redistributed Logged` +
                   `Change in Population` + `Municipality Area, Post-War Logged` +
                   `Sarajevo District` + `Croat Population Percentage, 1991` +
                   `Change in % Croats, 1991 to 2013`,
                 data = df_rs, family = gaussian())
summary(model_a2r)
car::vif(model_a2r)

# Federation
model_a2f <- glm(`Proportion of Votes for Non-Nationalist Parties, 1997` ~
                   `Deaths per Population` + `Minefields per Sq. km, Canton` + `Ethnic Fractionalization, 1991` +
                   `Change in Ethnic Fractionalization, 1991 to 2013` +
                   `Proportion of Votes Cast Out-District, 1997` + `Turnout Rate, 1997` +
                   `Yugoslav Population Percentage, 1991` + `Other Population Percentage, 1991` +
                   `Absolute Change in Bosniaks, Croats, and Serbs, 1991 to 2013` +
                   `Distance to Croatia, Post-War` + `Distance to Yugoslavia, Post-War` +
                   `Distance to IEBL, Post-War` + `Population, 1991 Redistributed Logged` +
                   `Change in Population` + `Municipality Area, Post-War Logged` +
                   `Sarajevo District` + `Croat Population Percentage, 1991` +
                   `Change in % Croats, 1991 to 2013`,
                 data = df_fbih, family = gaussian())
summary(model_a2f)
car::vif(model_a2f)

```

### A3 Models
```{r}
#| label: a3
#| echo: FALSE

# all municipalities
model_a3 <- glm(`Proportion of Votes for Non-Nationalist Parties, 1997` ~
                  `Deaths per Population` + `Minefields per Sq. km, Canton` + `Ethnic Fractionalization, 1991` +
                  `Change in Ethnic Fractionalization, 1991 to 2013` +
                  `Proportion of Votes Cast Out-District, 1997` + `Turnout Rate, 1997` +
                  `Yugoslav Population Percentage, 1991` + `Other Population Percentage, 1991` +
                  `Absolute Change in Bosniaks, Croats, and Serbs, 1991 to 2013` +
                  `Distance to Croatia, Post-War` + `Distance to Yugoslavia, Post-War` +
                  `Distance to IEBL, Post-War` + `Population, 1991 Redistributed Logged` +
                  `Change in Population` + `Municipality Area, Post-War Logged` +
                  `Federation Municipality` + `Sarajevo District` +
                  `Serb Population Percentage, 1991` + `Change in % Serbs, 1991 to 2013`,
                data = df, family = gaussian())
summary(model_a3)
car::vif(model_a3)

# Republika Srpska
model_a3r <- glm(`Proportion of Votes for Non-Nationalist Parties, 1997` ~
                   `Deaths per Population` + `Ethnic Fractionalization, 1991` +
                   `Change in Ethnic Fractionalization, 1991 to 2013` +
                   `Proportion of Votes Cast Out-District, 1997` + `Turnout Rate, 1997` +
                   `Yugoslav Population Percentage, 1991` + `Other Population Percentage, 1991` +
                   `Absolute Change in Bosniaks, Croats, and Serbs, 1991 to 2013` +
                   `Distance to Croatia, Post-War` + `Distance to Yugoslavia, Post-War` +
                   `Distance to IEBL, Post-War` + `Population, 1991 Redistributed Logged` +
                   `Change in Population` + `Municipality Area, Post-War Logged` +
                   `Sarajevo District` + `Serb Population Percentage, 1991` +
                   `Change in % Serbs, 1991 to 2013`,
                 data = df_rs, family = gaussian())
summary(model_a3r)
car::vif(model_a3r)

# Federation
model_a3f <- glm(`Proportion of Votes for Non-Nationalist Parties, 1997` ~
                   `Deaths per Population` + `Minefields per Sq. km, Canton` + `Ethnic Fractionalization, 1991` +
                   `Change in Ethnic Fractionalization, 1991 to 2013` +
                   `Proportion of Votes Cast Out-District, 1997` + `Turnout Rate, 1997` +
                   `Yugoslav Population Percentage, 1991` + `Other Population Percentage, 1991` +
                   `Absolute Change in Bosniaks, Croats, and Serbs, 1991 to 2013` +
                   `Distance to Croatia, Post-War` + `Distance to Yugoslavia, Post-War` +
                   `Distance to IEBL, Post-War` + `Population, 1991 Redistributed Logged` +
                   `Change in Population` + `Municipality Area, Post-War Logged` +
                   `Sarajevo District` + `Serb Population Percentage, 1991` +
                   `Change in % Serbs, 1991 to 2013`,
                 data = df_fbih, family = gaussian())
summary(model_a3f)
car::vif(model_a3f)

```

### A4 Models

### A5 Models

### A6 Models

## Nationalist Vote Share GLMs
* B1 (b1) - Country-wide / Bosniak / Fractionalization
* B1-RS (b1r) - Republika Srpska / Bosniak / Fractionalization
* B1-FBiH (b1f) - Federation / Bosniak / Fractionalization

* B2 (b2) - Country-wide / Croat / Fractionalization
* B2-RS (b2r) - Republika Srpska / Croat / Fractionalization
* B2-FBiH (b2f) - Federation / Croat / Fractionalization

* B3 (b3) - Country-wide / Serb / Fractionalization
* B3-RS (b3r) - Republika Srpska / Serb / Fractionalization
* B3-FBiH (b3f) - Federation / Serb / Fractionalization

* B4 (b4) - Country-wide / Bosniak / Polarization
* B4-RS (b4r) - Republika Srpska / Bosniak / Polarization
* B4-FBiH (b4f) - Federation / Bosniak / Polarization

* B5 (b5) - Country-wide / Croat / Polarization
* B5-RS (b5r) - Republika Srpska / Croat / Polarization
* B5-FBiH (b5f) - Federation / Croat / Polarization

* B6 (b6) - Country-wide / Serb / Polarization
* B6-RS (b6r) - Republika Srpska / Serb / Polarization
* B6-FBiH (b6f) - Federation / Serb / Polarization

### B1 models
```{r}
#| label: b1
#| echo: FALSE

# country-wide
model_b1 <- glm(`Proportion of Votes for Nationalist Parties, 1997` ~
                  `Deaths per Population` + `Minefields per Sq. km, Canton` + `Ethnic Fractionalization, 1991` +
                  `Change in Ethnic Fractionalization, 1991 to 2013` +
                  `Proportion of Votes Cast Out-District, 1997` + `Turnout Rate, 1997` +
                  `Yugoslav Population Percentage, 1991` + `Other Population Percentage, 1991` +
                  `Absolute Change in Bosniaks, Croats, and Serbs, 1991 to 2013` +
                  `Distance to Croatia, Post-War` + `Distance to Yugoslavia, Post-War` +
                  `Distance to IEBL, Post-War` + `Population, 1991 Redistributed Logged` +
                  `Change in Population` + `Municipality Area, Post-War Logged` +
                  `Federation Municipality` + `Sarajevo District` +
                  `Muslim Population Percentage, 1991` + `Change in % Bosniaks, 1991 to 2013`,
                data = df, family = gaussian())
summary(model_b1)
car::vif(model_b1)

# republika srpska
model_b1r <- glm(`Proportion of Votes for Nationalist Parties, 1997` ~
                   `Deaths per Population` + `Ethnic Fractionalization, 1991` +
                   `Change in Ethnic Fractionalization, 1991 to 2013` +
                   `Proportion of Votes Cast Out-District, 1997` + `Turnout Rate, 1997` +
                   `Yugoslav Population Percentage, 1991` + `Other Population Percentage, 1991` +
                   `Absolute Change in Bosniaks, Croats, and Serbs, 1991 to 2013` +
                   `Distance to Croatia, Post-War` + `Distance to Yugoslavia, Post-War` +
                   `Distance to IEBL, Post-War` + `Population, 1991 Redistributed Logged` +
                   `Change in Population` + `Municipality Area, Post-War Logged` +
                   `Sarajevo District` + `Muslim Population Percentage, 1991` +
                   `Change in % Bosniaks, 1991 to 2013`,
                 data = df_rs, family = gaussian())
summary(model_b1r)
car::vif(model_b1r)

# federation
model_b1f <- glm(`Proportion of Votes for Nationalist Parties, 1997` ~
                   `Deaths per Population` + `Minefields per Sq. km, Canton` + `Ethnic Fractionalization, 1991` +
                   `Change in Ethnic Fractionalization, 1991 to 2013` +
                   `Proportion of Votes Cast Out-District, 1997` + `Turnout Rate, 1997` +
                   `Yugoslav Population Percentage, 1991` + `Other Population Percentage, 1991` +
                   `Absolute Change in Bosniaks, Croats, and Serbs, 1991 to 2013` +
                   `Distance to Croatia, Post-War` + `Distance to Yugoslavia, Post-War` +
                   `Distance to IEBL, Post-War` + `Population, 1991 Redistributed Logged` +
                   `Change in Population` + `Municipality Area, Post-War Logged` +
                   `Sarajevo District` + `Muslim Population Percentage, 1991` +
                   `Change in % Bosniaks, 1991 to 2013`,
                 data = df_fbih, family = gaussian())
summary(model_b1f)
car::vif(model_b1f)

```

```{r}
#| label: b2
#| echo: FALSE

# republic wide
model_b2 <- glm(`Proportion of Votes for Nationalist Parties, 1997` ~
                  `Deaths per Population` + `Minefields per Sq. km, Canton` + `Ethnic Fractionalization, 1991` +
                  `Change in Ethnic Fractionalization, 1991 to 2013` +
                  `Proportion of Votes Cast Out-District, 1997` + `Turnout Rate, 1997` +
                  `Yugoslav Population Percentage, 1991` + `Other Population Percentage, 1991` +
                  `Absolute Change in Bosniaks, Croats, and Serbs, 1991 to 2013` +
                  `Distance to Croatia, Post-War` + `Distance to Yugoslavia, Post-War` +
                  `Distance to IEBL, Post-War` + `Population, 1991 Redistributed Logged` +
                  `Change in Population` + `Municipality Area, Post-War Logged` +
                  `Federation Municipality` + `Sarajevo District` +
                  `Croat Population Percentage, 1991` + `Change in % Croats, 1991 to 2013`,
                data = df, family = gaussian())
summary(model_b2)
car::vif(model_b2)

# republika srpska
model_b2r <- glm(`Proportion of Votes for Nationalist Parties, 1997` ~
                   `Deaths per Population` + `Ethnic Fractionalization, 1991` +
                   `Change in Ethnic Fractionalization, 1991 to 2013` +
                   `Proportion of Votes Cast Out-District, 1997` + `Turnout Rate, 1997` +
                   `Yugoslav Population Percentage, 1991` + `Other Population Percentage, 1991` +
                   `Absolute Change in Bosniaks, Croats, and Serbs, 1991 to 2013` +
                   `Distance to Croatia, Post-War` + `Distance to Yugoslavia, Post-War` +
                   `Distance to IEBL, Post-War` + `Population, 1991 Redistributed Logged` +
                   `Change in Population` + `Municipality Area, Post-War Logged` +
                   `Sarajevo District` + `Croat Population Percentage, 1991` +
                   `Change in % Croats, 1991 to 2013`,
                 data = df_rs, family = gaussian())
summary(model_b2r)
car::vif(model_b2r)

# federation
model_b2f <- glm(`Proportion of Votes for Nationalist Parties, 1997` ~
                   `Deaths per Population` + `Minefields per Sq. km, Canton` + `Ethnic Fractionalization, 1991` +
                   `Change in Ethnic Fractionalization, 1991 to 2013` +
                   `Proportion of Votes Cast Out-District, 1997` + `Turnout Rate, 1997` +
                   `Yugoslav Population Percentage, 1991` + `Other Population Percentage, 1991` +
                   `Absolute Change in Bosniaks, Croats, and Serbs, 1991 to 2013` +
                   `Distance to Croatia, Post-War` + `Distance to Yugoslavia, Post-War` +
                   `Distance to IEBL, Post-War` + `Population, 1991 Redistributed Logged` +
                   `Change in Population` + `Municipality Area, Post-War Logged` +
                   `Sarajevo District` + `Croat Population Percentage, 1991` +
                   `Change in % Croats, 1991 to 2013`,
                 data = df_fbih, family = gaussian())
summary(model_b2f)
car::vif(model_b2f)
```

### B3 models

```{r}
#| label: b3
#| echo: FALSE

# republic wide
model_b3 <- glm(`Proportion of Votes for Nationalist Parties, 1997` ~
                  `Deaths per Population` + `Minefields per Sq. km, Canton` + `Ethnic Fractionalization, 1991` +
                  `Change in Ethnic Fractionalization, 1991 to 2013` +
                  `Proportion of Votes Cast Out-District, 1997` + `Turnout Rate, 1997` +
                  `Yugoslav Population Percentage, 1991` + `Other Population Percentage, 1991` +
                  `Absolute Change in Bosniaks, Croats, and Serbs, 1991 to 2013` +
                  `Distance to Croatia, Post-War` + `Distance to Yugoslavia, Post-War` +
                  `Distance to IEBL, Post-War` + `Population, 1991 Redistributed Logged` +
                  `Change in Population` + `Municipality Area, Post-War Logged` +
                  `Federation Municipality` + `Sarajevo District` +
                  `Serb Population Percentage, 1991` + `Change in % Serbs, 1991 to 2013`,
                data = df, family = gaussian())
summary(model_b3)
car::vif(model_b3)

# republika srpska
model_b3r <- glm(`Proportion of Votes for Nationalist Parties, 1997` ~
                   `Deaths per Population` + `Ethnic Fractionalization, 1991` +
                   `Change in Ethnic Fractionalization, 1991 to 2013` +
                   `Proportion of Votes Cast Out-District, 1997` + `Turnout Rate, 1997` +
                   `Yugoslav Population Percentage, 1991` + `Other Population Percentage, 1991` +
                   `Absolute Change in Bosniaks, Croats, and Serbs, 1991 to 2013` +
                   `Distance to Croatia, Post-War` + `Distance to Yugoslavia, Post-War` +
                   `Distance to IEBL, Post-War` + `Population, 1991 Redistributed Logged` +
                   `Change in Population` + `Municipality Area, Post-War Logged` +
                   `Sarajevo District` + `Serb Population Percentage, 1991` +
                   `Change in % Serbs, 1991 to 2013`,
                 data = df_rs, family = gaussian())
summary(model_b3r)
car::vif(model_b3r)

# federation
model_b3f <- glm(`Proportion of Votes for Nationalist Parties, 1997` ~
                   `Deaths per Population` + `Minefields per Sq. km, Canton` + `Ethnic Fractionalization, 1991` +
                   `Change in Ethnic Fractionalization, 1991 to 2013` +
                   `Proportion of Votes Cast Out-District, 1997` + `Turnout Rate, 1997` +
                   `Yugoslav Population Percentage, 1991` + `Other Population Percentage, 1991` +
                   `Absolute Change in Bosniaks, Croats, and Serbs, 1991 to 2013` +
                   `Distance to Croatia, Post-War` + `Distance to Yugoslavia, Post-War` +
                   `Distance to IEBL, Post-War` + `Population, 1991 Redistributed Logged` +
                   `Change in Population` + `Municipality Area, Post-War Logged` +
                   `Sarajevo District` + `Serb Population Percentage, 1991` +
                   `Change in % Serbs, 1991 to 2013`,
                 data = df_fbih, family = gaussian())
summary(model_b3f)
car::vif(model_b3f)

```

## Model Summaries
```{r}

# A - Entire Country
modelsummary::modelsummary(
  list("A1" = model_a1, "A2" = model_a2, "A3" = model_a3),
  output = "markdown",
  stars = TRUE
  )
# A - Republika Srpska
modelsummary::modelsummary(
  list("A1-RS" = model_a1r, "A2-RS" = model_a2r, "A3-RS" = model_a3r),
  output = "markdown",
  stars = TRUE
  )
# A - Federation
modelsummary::modelsummary(
  list("A1-FBiH" = model_a1f, "A2-FBiH" = model_a2f, "A3-FBiH" = model_a3f),
  output = "markdown",
  stars = TRUE
  )

# B - Entire Country
modelsummary::modelsummary(
  list("B1" = model_b1, "B2" = model_b2, "B3" = model_b3),
  output = "markdown",
  stars = TRUE
  )
# B - Republika Srpska
modelsummary::modelsummary(
  list("B1-RS" = model_b1r, "B2-RS" = model_b2r, "B3-RS" = model_b3r),
  output = "markdown",
  stars = TRUE
  )
# B - Federation
modelsummary::modelsummary(
  list("B1-FBiH" = model_b1f, "B2-FBiH" = model_b2f, "B3-FBiH" = model_b3f),
  output = "markdown",
  stars = TRUE
  )

```

